Nanophotonics with optical phonons: Phonon-polaritonic response without $\Gamma$-point LO-TO splitting in two-dimensional materials

### 2019-01-03 paper.tex
starting on the edits..

% --- TITLE & AUTHORSHIP ---
\title{Phonon-polaritonic reponse without $\Gamma$-point LO-TO splitting in two-dimensional materials\\ or \\ Nanophotonics with optical phonons in two dimensions}

@@

% --- ABSTRACT ---
\begin{abstract}
    Extreme confinement of electromagnetic energy by phonon-polaritons holds the promise of extremely strong and novel forms of control over the dynamics of matter. 
    To bring such control to its ultimate limit, it is important to consider phonon-polaritons in two dimensional systems.  Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.  Does this lack of LO--TO splitting imply the absence of a phonon-polariton in polar monolayers? 
    Here, we derive an expression for  the spatially non-local conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions. In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the TO phonon frequency at the $\Gamma$ point, its lifetime, and the group velocity of the LO phonon. Our analysis reveals that the phonon-polariton of the bulk is essentially replaced by the LO phonon of the two-dimensional system. We discuss the confinement and propagation losses of the LO phonons, finding that high  confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques. 
    We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
\end{abstract}

>>>>

% --- TITLE & AUTHORSHIP ---
\title{Nanophotonics with optical phonons: Phonon-polaritonic response without $\Gamma$-point LO-TO splitting in two-dimensional materials}

@@

% --- ABSTRACT ---
\begin{abstract}
Extreme confinement of electromagnetic energy by phonon-polaritons holds the promise of extremely strong and novel forms of control over the dynamics of matter. To bring such control to the atomic-scale limit, it is important to consider phonon-polaritons in two dimensional systems.  Recent studies have pointed out that in two-dimensional systems, splitting between longitudinal and transverse optical (LO and TO) phonons is absent at the $\Gamma$ point.  Does this lack of LO--TO splitting imply the absence of a phonon-polariton in polar monolayers? 
    Here, we derive an expression for  the spatially non-local conductivity of a polar monolayer specified by the wavevector-dependent LO and TO phonon dispersions. In the long-wavelength (local) limit, we find a universal form for the conductivity in terms of the TO phonon frequency at the $\Gamma$ point, its lifetime, and the group velocity of the LO phonon. Our analysis reveals that the phonon-polariton of the bulk is essentially replaced by the LO phonon of the two-dimensional system. We discuss the confinement and propagation losses of the LO phonons, finding that high  confinement and reasonable propagation quality factors coincide in regions which are difficult to detect with current near-field optical microscope techniques. 
    We conclude by studying the interaction of external emitters with two-dimensional hBN nanostructures, finding extreme enhancement of spontaneous emission due to coupling with localized 2D LO phonons, and the possibility of multi-mode strong and ultra-strong coupling between an external emitter and hBN phonons.
\end{abstract}
